Smart Drill-Down

### 2014-11-30 TableSummarization/Paper/algorithms.tex
Replaced drill-down with drill down everywhere

\section{Smart Drill-Down Algorithms} \label{sec:algorithms}
We now describe online algorithms for implementing
the smart drill-down operator. We assume in this section that all columns are categorical (so numerical columns have been bucketized beforehand). We further discuss bucketization of numerical attributes in the Extensions section in the technical report~\cite{tr}.

@@

\noindent Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
\begin{enumerate}
\item $[$Rule Drill-Down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$. Then, we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get a smaller table $T_r$ consisting of tuples from $T$ that are covered by $r$. In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise. Then, we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
\end{enumerate}

>>>>

\section{Smart drill down Algorithms} \label{sec:algorithms}
We now describe online algorithms for implementing
the smart drill down operator. We assume in this section that all columns are categorical (so numerical columns have been bucketized beforehand). We further discuss bucketization of numerical attributes in the Extensions section in the technical report~\cite{tr}.

@@

\noindent Problem~\ref{prob:optimal-subrule-list} with parameters $(T, W, k)$ can be reduced to Problem~\ref{prob:optimal-rule-list} as follows:
\begin{enumerate}
\item $[$Rule drill down$]$ If the user clicked on rule $r$ in Problem~\ref{prob:optimal-subrule-list}, then we can conceptually make one pass through the table $T$ to filter for tuples covered by rule $r$, and store them in a temporary table $T_r$. Then, we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W, k)$.
\item $[$Star drill down$]$ If the user clicked on a $\star$ in column $c$ of rule $r$, then we first filter table $T$ to get a smaller table $T_r$ consisting of tuples from $T$ that are covered by $r$. In addition, we change the weight function $W$ from Problem~\ref{prob:optimal-subrule-list} to a weight function $W^{\prime}$ such that : For any rule $r^{\prime}$, $W^{\prime}(r^{\prime}) = 0$ if $r^{\prime}$ has a $\star$ in column $c$, and $W^{\prime}(r^{\prime}) = W(r^{\prime})$ otherwise. Then, we solve Problem~\ref{prob:optimal-rule-list} for parameters $(T_r, W^{\prime}, k)$.
\end{enumerate}

### 2014-11-30 TableSummarization/Paper/exp.tex
Replaced drill-down with drill down everywhere

\section{Experiments}\label{sec:experiments}
We have implemented a fully-functional interactive system instrumented with the smart drill-down operator.
We now describe our experiments on this system with real datasets.

@@

Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit in the page. We use our current implementation of our the smart drill-down operator, which can directly output tables in \LaTeX \ format. We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting and $20$ for the Bits weighting function (see Section~\ref{sec:weighting}). Memory capacity $M$ for the SampleHandler is set to $50000$ tuples, and $minSS$ to $5000$.

\subsection{Qualitative Study}
We first perform a qualitative study of smart drill-down. We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out a few different weight functions to study their effects.
\subsubsection{User interface testing}
We now present the rule-based summaries displayed as a result of a few different user actions.

To begin with, the user sees an empty rule with the total number of tuples as the count. Suppose the user expands the rule. Then the user will see Table~\ref{table:uiexample1}. The first two new rules simply tell us that the table has $4918$ female and $4075$ male tuples. The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Area for $> 10$ years. Note that the latter two rules give very specific information which would require up to $3$ user clicks to find using traditional drill-down, whereas smart drill-down displays that information to the user with a single click.

>>>>

\section{Experiments}\label{sec:experiments}
We have implemented a fully-functional interactive system instrumented with the smart drill down operator.
We now describe our experiments on this system with real datasets.

@@

Unless otherwise specified, in all our experiments, we restrict the tables to the first $7$ columns in order to make the result tables fit in the page. We use our current implementation of our the smart drill down operator, which can directly output tables in \LaTeX \ format. We set the $k$ (number of rules) parameter to $4$, and $m_w$ to $5$ for the Size weighting and $20$ for the Bits weighting function (see Section~\ref{sec:weighting}). Memory capacity $M$ for the SampleHandler is set to $50000$ tuples, and $minSS$ to $5000$.

\subsection{Qualitative Study}
We first perform a qualitative study of smart drill down. We observe the effects of various user interface operations on the Marketing Dataset (the results are similar on the Census dataset), and then try out a few different weight functions to study their effects.
\subsubsection{User interface testing}
We now present the rule-based summaries displayed as a result of a few different user actions.

To begin with, the user sees an empty rule with the total number of tuples as the count. Suppose the user expands the rule. Then the user will see Table~\ref{table:uiexample1}. The first two new rules simply tell us that the table has $4918$ female and $4075$ male tuples. The next two rules also slightly more detailed, saying that there are $2940$ females who have been in the Bay Area for $> 10$ years, and $980$ males who have never been married and been in the Bay Area for $> 10$ years. Note that the latter two rules give very specific information which would require up to $3$ user clicks to find using traditional drill down, whereas smart drill down displays that information to the user with a single click.

### 2014-11-30 TableSummarization/Paper/formal.tex
Replaced drill-down with drill down everywhere

Our objective (formally defined later) is to 
enable smart drill-downs on this table or on portions of it:
the result of our drill-downs are lists of {\em rules}. 
A {\em rule} is a tuple with a value for each column of the table. 
In addition, a rule has other attributes, such as count and weight 
(which we define later) associated with it. 
The value in each column of the rule can either be one of the values in the corresponding column of the table, or $\star$, representing a wildcard character representing all values in the column. For a column with numerical values in the table, we allow the corresponding rule-value to be a range instead of a single value. The {\em trivial rule} is one that has a $\star$ value in all columns. The {\em Size} of a rule is defined as the number of non-starred values in that rule.

@@

\stitle{Rule Lists:}
A {\em rule-list} is an ordered list of rules returned by our system when a smart drill-down operation concludes. 
When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a rule-list of sub-rules of $r$.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user clicks on the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

@@

We now describe the first portion:
we define {\em Count}($r$) as the total number of tuples $t \in T$ that are covered by $r$. 
Further, we define {\em MCount}($r, R$) (which stands for `Marginal Count') as the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. A high value of $MCount$ indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules. We want to pick rules with a high value of $MCount$ to display to the user
as part of the smart drill-down result, to increase the coverage of the rule-list.

@@

Our smart drill-downs still display the Count of each rule rather than the $MCount$. This is because while $MCount$ is useful in the rule selection process, Count is easier for a user to interpret. In any case, it would be a simple extension to display MCount in another column.

\stitle{Formal Problem:} We now formally define our problem:
\begin{problem}\label{prob:optimal-subrule-list}
Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r) \times MCount(r,R)$$
for one of the following smart drill-down operations:
\begin{denselist}
\item $[$Rule Drill-Down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item $[$Star Drill-Down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{denselist}
\end{problem}
\noindent Throughout this paper, we use the {\em Count} aggregate of a rule to display to the user. We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead. We discuss how to modify our algorithms to use $Sum$ instead of $Count$ 
\techreporttext{in the `Extensions' section}
\papertext{in the technical report~\cite{tr}}.

@@

\subsection{Smart Drill-Down Operations}
\label{sec:interface}
When the user starts using a system equipped with
the smart drill-down operator, they first see a table with a single trivial rule as shown in table~\ref{table:introexample0}. At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill-down' on the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule of table~\ref{table:introexample0} would display table~\ref{table:introexample}. Clicking further on the third rule in the expanded rule-list would display table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

@@

Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e. collapses it. For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}. This operation is equivalent to a traditional
roll-up, but for smart drill-downs instead of traditional drill-downs.

>>>>

Our objective (formally defined later) is to 
enable smart drill downs on this table or on portions of it:
the result of our drill downs are lists of {\em rules}. 
A {\em rule} is a tuple with a value for each column of the table. 
In addition, a rule has other attributes, such as count and weight 
(which we define later) associated with it. 
The value in each column of the rule can either be one of the values in the corresponding column of the table, or $\star$, representing a wildcard character representing all values in the column. For a column with numerical values in the table, we allow the corresponding rule-value to be a range instead of a single value. The {\em trivial rule} is one that has a $\star$ value in all columns. The {\em Size} of a rule is defined as the number of non-starred values in that rule.

@@

\stitle{Rule Lists:}
A {\em rule-list} is an ordered list of rules returned by our system when a smart drill down operation concludes. 
When a user drills down on a rule $r$ to know more about the part of the table covered by $r$, we display a rule-list of sub-rules of $r$.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user clicks on the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

@@

We now describe the first portion:
we define {\em Count}($r$) as the total number of tuples $t \in T$ that are covered by $r$. 
Further, we define {\em MCount}($r, R$) (which stands for `Marginal Count') as the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. A high value of $MCount$ indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules. We want to pick rules with a high value of $MCount$ to display to the user
as part of the smart drill down result, to increase the coverage of the rule-list.

@@

Our smart drill downs still display the Count of each rule rather than the $MCount$. This is because while $MCount$ is useful in the rule selection process, Count is easier for a user to interpret. In any case, it would be a simple extension to display MCount in another column.

\stitle{Formal Problem:} We now formally define our problem:
\begin{problem}\label{prob:optimal-subrule-list}
Given a table $T$, a monotonic weighting function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes 
$$\sum_{r \in R} W(r) \times MCount(r,R)$$
for one of the following smart drill down operations:
\begin{denselist}
\item $[$Rule drill down$]$ If the user clicked on a rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$
\item $[$Star drill down$]$ If the user clicked on a $\star$ on column $c$ of rule $r^{\prime}$, then all $r \in R$ must be super-rules of $r^{\prime}$ and have a non-$\star$ value in column $c$
\end{denselist}
\end{problem}
\noindent Throughout this paper, we use the {\em Count} aggregate of a rule to display to the user. We can also use a {\em Sum} of values over a given `measure column' $c_m$ instead. We discuss how to modify our algorithms to use $Sum$ instead of $Count$ 
\techreporttext{in the `Extensions' section}
\papertext{in the technical report~\cite{tr}}.

@@

\subsection{Smart drill down Operations}
\label{sec:interface}
When the user starts using a system equipped with
the smart drill down operator, they first see a table with a single trivial rule as shown in table~\ref{table:introexample0}. At any point, the user can click on either a rule, or a star within a rule, to perform a `smart drill down' on the rule. Clicking on a rule $r$ causes $r$ to expand into the highest-scoring rule-list consisting of super-rules of $r$. By default, the rule $r$ expands into a list of $3$ rules, but this number can be changed by the user. As an example of this operation, clicking on the trivial rule of table~\ref{table:introexample0} would display table~\ref{table:introexample}. Clicking further on the third rule in the expanded rule-list would display table~\ref{table:introexample2}. The rules obtained from the expansion are listed directly below $r$, ordered in decreasing order by weight (the reasoning behind the ordering is explained in Section~\ref{sec:algorithms}).

@@

Finally, when the user clicks on a rule that has already been expanded, it reverses the expansion operation, i.e. collapses it. For example, clicking on the walmart rule in table~\ref{table:introexample3} or table~\ref{table:introexample2} would take the user back to table~\ref{table:introexample}. This operation is equivalent to a traditional
roll-up, but for smart drill downs instead of traditional drill downs.

### 2014-11-30 TableSummarization/Paper/intro.tex
Replaced drill-down with drill down everywhere

\section{Introduction}
Analysts often use OLAP (Online Analytical Processing) operations
such as drill-down (and roll-up)~\cite{export:69578} to explore
relational databases. 
These operations are very useful for analytics and data exploration and have stood the test of time;
all commercial OLAP systems 
in existence support these operations. (Recent reports estimate the size of the OLAP market to be \$10+ Billion~\cite{gartner}.)

However, there are cases where drill-down is ineffective; 
for example, when the number of distinct values
in a column is large, vanilla drill-down 
could easily overwhelm analysts by presenting them with too many 
results (i.e., aggregates). 
Further, drill-down only allows us to instantiate values
 one column at a time, instead of allowing simultaneous drill-downs
on multiple columns---this simultaneous drill-down on multiple columns 
could once again suffer from the problem
of having too many results, stemming from many distinct combinations of column values.

In this paper, we present a new interaction operator 
that is an extension to a traditional 
drill-down operator, aimed at providing {\em complementary}
functionality to drill-down in cases where drill-down is
ineffective. We call our operator {\em smart drill-down}.
At a high level, smart drill-down lets analysts zoom into
the more ``interesting'' parts of a table or a database,
with fewer operations, and without having to examine as much
data as traditional drill-down.
Note that our goal is {\em not} to replace traditional 
drill-down functionality, which we believe is fundamental;
instead, our goal is to provide auxiliary functionality 
which the analyst is free to use whenever they find 
traditional drill-downs ineffective.

In addition to presenting this new operator called smart drill-down, in this paper, 
we also present novel sampling techniques to compute the results for this
operator {\em in an interactive fashion} on increasingly larger databases. 
Unlike the traditional OLAP setting, these computations 
require no pre-materialization, and can be implemented 
within or on top of any relational database system.

% drill-down (and roll up)~\cite{export:69578} are very useful tools for exploring
% relational tables.
% In this paper we study an extension of drill-down
% that lets users zoom into the more ``interesting'' parts of a table,
% with fewer operations and without having to examine as much data
% as traditional drill-down.
% We call our approach {\em smart drill-down};
% it does not replace traditional drill-down but we will argue is a very
% useful complement.
The best way to explain smart drill-down is through a simple example.

\begin{example}\label{ex:introexample}
Consider a table with columns `Department Store', `Product', `Region'
and `Sales'. Suppose an analyst queries for tuples
where Sales were higher than some threshold, in order
to find the best selling products.
If the resulting table has many tuples,
the analyst can use traditional drill-down to explore it.
For instance, the system may initially tell the analyst there are
6000 tuples in the answer, represented by the tuple ($\star$, $\star$, $\star$, $6000$, $0$),
as shown in Table~\ref{table:introexample0}.
The $\star$ character is a wildcard that matches any value in the database.
The Count attribute can be replaced by a Sum aggregate over some measure column,
e.g., the total sales.
The right-most Weight attribute is the number of non-$\star$ attributes; 
its significance will be discussed shortly.
If the analyst drills down on the Store attribute (first $\star$),
then the operator displays all tuples of the form ($X$, $\star$, $\star$, $C$, $1$),
where $X$ is a Store in the answer table, and $C$
is the number of tuples for $X$ (or the aggregate sales for $X$).

Instead, when the analyst uses smart drill-down on Table~\ref{table:introexample0},
he obtains Table~\ref{table:introexample}.
The ($\star$, $\star$, $\star$, $6000$) tuple is expanded into $3$ tuples
that display noteworthy or interesting drill-downs.
The number $3$ is a user specified parameter, which we call $k$.

@@

\begin{table}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \hline
$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \hline
$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\vspace{-10pt}
\caption{Result after first smart drill-down}\label{table:introexample}
\vspace{-10pt}
\end{table}

\begin{table}
\scriptsize
\centering
\vspace{-5pt}
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & $\star$ & CA-1 & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & $\star$ & WA-5 & $130$ & $2$ \\ \hline
\end{tabular}
\vspace{-10pt}
\caption{Result after second smart drill-down} \label{table:introexample2}
\vspace{-10pt}
\end{table}

@@

When the analyst clicks on a rule $r$, smart drill-down
expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
(We discuss other smart drill-down operations in Section~\ref{sec:interface}.)
There are three factors that make a rule set interesting.
One is if it contains rules with high Count (or total sales) fields,
since the larger the count, the more tuples are summarized.
A second factor is if the rules have high weight (number of non-$\star$ attributes).
For instance, the rule (Walmart, cookies, AK-1, $200$, $3$)
seems more interesting than (Walmart, cookies, $*$, $200$, $2$)
since the former tells us the high sales are concentrated in a single region.
A third desirability factor is diversity:
For example, if we already have the rule (Walmart, $\star$, $\star$, $1000$, $1$)
in our set, we would rather have the rule (Target, bicycles, $\star$, $200$, $2$)
than (Walmart, bicycles, $\star$, $200$, $2$) since the former rule
describes tuples that are not described by the first rule.

In this paper we describe how to combine or blend these three factors
in order to obtain a single desirability score for a set of rules.
Our score function can actually be tuned by the analyst
(by specifying how weights are computed),
providing significant flexibility in what is considered a good set of rules.
We also present an efficient optimization procedure to maximize score, invoked
by smart drill-down to select the set of $k$ rules to display.

@@

Compared to traditional drill-down, our smart drill-down has two important advantages:
\squishlist
\item
Smart drill-down limits the information displayed
to the most interesting $k$ facts (rules).
With traditional drill-down, a column is expanded and {\em all}
attribute values are displayed in arbitrary order.
In our example, if we drill-down on say the store attribute,
we would see all stores listed, which may be a very large number.
\item
Smart drill-down explores several attributes to open up together,
and automatically selects combinations that are interesting.
For example, in Table~\ref{table:introexample},
the rule (Target, bicycles, $\star$, $200$, $2$)
is obtained after a single drill-down;
with a traditional approach, the analyst would first have to drill-down on
Store, examine the results, drill-down on Product,
look through all the displayed rules and then find the interesting rule
(Target, bicycles, $\star$, $200$, $2$).
\squishend

Incidentally, note that in the example we only described one type of smart drill-down,
where the analyst selects a {\em rule} to drill-down on
(e.g., the Walmart rule going from Table~\ref{table:introexample} to
Table~\ref{table:introexample2}.
In Section~\ref{sec:interface} we describe another option
where the analyst clicks on a $\star$ in a column to obtain
rules that have non-$\star$ values in that column.

Our work on smart drill-down is related
to table summarization and anomaly
detection~\cite{Sarawagi:2001:UMA:767141.767148,
Sarawagi00user-adaptiveexploration,
Sarawagi98discovery-drivenexploration,
DBLP:journals/pvldb/GebalyAGKS14}.
These papers mostly focus on
giving the most ``surprising'' information to the user, i.e., information
that would minimize the Kullback-Liebler(KL) divergence between the
resulting maximum entropy distribution and the actual value distribution. For instance, if a certain set of
values occur together in an unexpectedly small number of tuples, that
set of values may be displayed to the user. In contrast, our algorithm
focuses on rules with high counts, covering as
much of the table as possible.
Furthermore, our summarization is couched in
an interactive environment, where the analyst
directs the drill-down and can tailor the optimization criteria.
Nevertheless, one can envision extending traditional
and smart drill-down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}.

To reiterate, our chief contribution in this paper is the {\em smart drill-down} interaction operator,
an extension of traditional drill-down, aimed at allowing analysts to zoom into the
more ``interesting'' parts of a dataset. In addition to this operator,
we develop techniques to support this operator on increasingly larger datasets:
\squishlist
\item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill-down operation. b
\item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
maintaining samples in memory to support smart drill-down. 
We show that optimal identification of samples is once again {\sc NP-Hard}, and 
we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.
% to improve response time on large tables.
\squishend

@@

\item In Section~\ref{sec:formal}, we formally define smart drill-down. After that, we describe different schemes for weighting rules, and our interactive user interface.

@@

\item Based on our implemented smart drill-down,
in Section~\ref{sec:experiments} we experimentally evaluate
performance on real datasets,
and show additional examples of smart drill-down in action.

>>>>

\section{Introduction}
Analysts often use OLAP (Online Analytical Processing) operations
such as drill down (and roll-up)~\cite{export:69578} to explore
relational databases. 
These operations are very useful for analytics and data exploration and have stood the test of time;
all commercial OLAP systems 
in existence support these operations. (Recent reports estimate the size of the OLAP market to be \$10+ Billion~\cite{gartner}.)

However, there are cases where drill down is ineffective; 
for example, when the number of distinct values
in a column is large, vanilla drill down 
could easily overwhelm analysts by presenting them with too many 
results (i.e., aggregates). 
Further, drill down only allows us to instantiate values
 one column at a time, instead of allowing simultaneous drill downs
on multiple columns---this simultaneous drill down on multiple columns 
could once again suffer from the problem
of having too many results, stemming from many distinct combinations of column values.

In this paper, we present a new interaction operator 
that is an extension to a traditional 
drill down operator, aimed at providing {\em complementary}
functionality to drill down in cases where drill down is
ineffective. We call our operator {\em smart drill down}.
At a high level, smart drill down lets analysts zoom into
the more ``interesting'' parts of a table or a database,
with fewer operations, and without having to examine as much
data as traditional drill down.
Note that our goal is {\em not} to replace traditional 
drill down functionality, which we believe is fundamental;
instead, our goal is to provide auxiliary functionality 
which the analyst is free to use whenever they find 
traditional drill downs ineffective.

In addition to presenting this new operator called smart drill down, in this paper, 
we also present novel sampling techniques to compute the results for this
operator {\em in an interactive fashion} on increasingly larger databases. 
Unlike the traditional OLAP setting, these computations 
require no pre-materialization, and can be implemented 
within or on top of any relational database system.

% drill down (and roll up)~\cite{export:69578} are very useful tools for exploring
% relational tables.
% In this paper we study an extension of drill down
% that lets users zoom into the more ``interesting'' parts of a table,
% with fewer operations and without having to examine as much data
% as traditional drill down.
% We call our approach {\em smart drill down};
% it does not replace traditional drill down but we will argue is a very
% useful complement.
The best way to explain smart drill down is through a simple example.

\begin{example}\label{ex:introexample}
Consider a table with columns `Department Store', `Product', `Region'
and `Sales'. Suppose an analyst queries for tuples
where Sales were higher than some threshold, in order
to find the best selling products.
If the resulting table has many tuples,
the analyst can use traditional drill down to explore it.
For instance, the system may initially tell the analyst there are
6000 tuples in the answer, represented by the tuple ($\star$, $\star$, $\star$, $6000$, $0$),
as shown in Table~\ref{table:introexample0}.
The $\star$ character is a wildcard that matches any value in the database.
The Count attribute can be replaced by a Sum aggregate over some measure column,
e.g., the total sales.
The right-most Weight attribute is the number of non-$\star$ attributes; 
its significance will be discussed shortly.
If the analyst drills down on the Store attribute (first $\star$),
then the operator displays all tuples of the form ($X$, $\star$, $\star$, $C$, $1$),
where $X$ is a Store in the answer table, and $C$
is the number of tuples for $X$ (or the aggregate sales for $X$).

Instead, when the analyst uses smart drill down on Table~\ref{table:introexample0},
he obtains Table~\ref{table:introexample}.
The ($\star$, $\star$, $\star$, $6000$) tuple is expanded into $3$ tuples
that display noteworthy or interesting drill downs.
The number $3$ is a user specified parameter, which we call $k$.

@@

\begin{table}
\scriptsize
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \hline
$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \hline
$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\vspace{-10pt}
\caption{Result after first smart drill down}\label{table:introexample}
\vspace{-10pt}
\end{table}

\begin{table}
\scriptsize
\centering
\vspace{-5pt}
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & Region & Count & Weight \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
$\triangleright$ Target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\triangleright$ $\star$ & comforters & MA-3 & $600$ & $2$ \\ \cline{1-5}
$\triangleright$ Walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & $\star$ & CA-1 & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ $\triangleright$ Walmart & $\star$ & WA-5 & $130$ & $2$ \\ \hline
\end{tabular}
\vspace{-10pt}
\caption{Result after second smart drill down} \label{table:introexample2}
\vspace{-10pt}
\end{table}

@@

When the analyst clicks on a rule $r$, smart drill down
expands $r$ into $k$ sub-rules that as a set are deemed to be ``interesting.''
(We discuss other smart drill down operations in Section~\ref{sec:interface}.)
There are three factors that make a rule set interesting.
One is if it contains rules with high Count (or total sales) fields,
since the larger the count, the more tuples are summarized.
A second factor is if the rules have high weight (number of non-$\star$ attributes).
For instance, the rule (Walmart, cookies, AK-1, $200$, $3$)
seems more interesting than (Walmart, cookies, $*$, $200$, $2$)
since the former tells us the high sales are concentrated in a single region.
A third desirability factor is diversity:
For example, if we already have the rule (Walmart, $\star$, $\star$, $1000$, $1$)
in our set, we would rather have the rule (Target, bicycles, $\star$, $200$, $2$)
than (Walmart, bicycles, $\star$, $200$, $2$) since the former rule
describes tuples that are not described by the first rule.

In this paper we describe how to combine or blend these three factors
in order to obtain a single desirability score for a set of rules.
Our score function can actually be tuned by the analyst
(by specifying how weights are computed),
providing significant flexibility in what is considered a good set of rules.
We also present an efficient optimization procedure to maximize score, invoked
by smart drill down to select the set of $k$ rules to display.

@@

Compared to traditional drill down, our smart drill down has two important advantages:
\squishlist
\item
Smart drill down limits the information displayed
to the most interesting $k$ facts (rules).
With traditional drill down, a column is expanded and {\em all}
attribute values are displayed in arbitrary order.
In our example, if we drill down on say the store attribute,
we would see all stores listed, which may be a very large number.
\item
Smart drill down explores several attributes to open up together,
and automatically selects combinations that are interesting.
For example, in Table~\ref{table:introexample},
the rule (Target, bicycles, $\star$, $200$, $2$)
is obtained after a single drill down;
with a traditional approach, the analyst would first have to drill down on
Store, examine the results, drill down on Product,
look through all the displayed rules and then find the interesting rule
(Target, bicycles, $\star$, $200$, $2$).
\squishend

Incidentally, note that in the example we only described one type of smart drill down,
where the analyst selects a {\em rule} to drill down on
(e.g., the Walmart rule going from Table~\ref{table:introexample} to
Table~\ref{table:introexample2}.
In Section~\ref{sec:interface} we describe another option
where the analyst clicks on a $\star$ in a column to obtain
rules that have non-$\star$ values in that column.

Our work on smart drill down is related
to table summarization and anomaly
detection~\cite{Sarawagi:2001:UMA:767141.767148,
Sarawagi00user-adaptiveexploration,
Sarawagi98discovery-drivenexploration,
DBLP:journals/pvldb/GebalyAGKS14}.
These papers mostly focus on
giving the most ``surprising'' information to the user, i.e., information
that would minimize the Kullback-Liebler(KL) divergence between the
resulting maximum entropy distribution and the actual value distribution. For instance, if a certain set of
values occur together in an unexpectedly small number of tuples, that
set of values may be displayed to the user. In contrast, our algorithm
focuses on rules with high counts, covering as
much of the table as possible.
Furthermore, our summarization is couched in
an interactive environment, where the analyst
directs the drill down and can tailor the optimization criteria.
Nevertheless, one can envision extending traditional
and smart drill down to provide an additional option of anomaly detection. We discuss related work in detail in  Section~\ref{sec:related}.

To reiterate, our chief contribution in this paper is the {\em smart drill down} interaction operator,
an extension of traditional drill down, aimed at allowing analysts to zoom into the
more ``interesting'' parts of a dataset. In addition to this operator,
we develop techniques to support this operator on increasingly larger datasets:
\squishlist
\item {\em Basic Interaction:} We demonstrate that finding the optimal list of rules is {\sc NP-Hard},
and we develop an algorithm to find the approximately optimal list of rules to display when the user performs a smart drill down operation. b
\item {\em Dynamic Sample Maintenance:} To improve response time on large tables, we formalize the problem of dynamically
maintaining samples in memory to support smart drill down. 
We show that optimal identification of samples is once again {\sc NP-Hard}, and 
we develop an approximate scheme for dynamically maintaining and using multiple samples of the table in memory.
% to improve response time on large tables.
\squishend

@@

\item In Section~\ref{sec:formal}, we formally define smart drill down. After that, we describe different schemes for weighting rules, and our interactive user interface.

@@

\item Based on our implemented smart drill down,
in Section~\ref{sec:experiments} we experimentally evaluate
performance on real datasets,
and show additional examples of smart drill down in action.

### 2014-11-30 TableSummarization/Paper/related.tex
Replaced drill-down with drill down everywhere

Some related work~\cite{Golab_efficientand, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies. There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, Xiang_succinctsummarization, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables. In both these cases, the emphasis is on completely covering or 
summarizing the table, suffering from the same problems as traditional drill-down in that the user may be presented with
too many results. The techniques in the former case may end up picking rare ``patterns'' if they have high confidence, and in the latter case only work for a restricted number of attributes (in their case, 4). \agp{check}

>>>>

Some related work~\cite{Golab_efficientand, Golab:2008:GNT:1453856.1453900} focuses on finding minimum sized Tableaux that provide improved support and confidence for conditional functional dependencies. There has also been work~\cite{Bu:2005:MSH:1083592.1083644, Lakshmanan:2002:GMA:1287369.1287435, Xiang_succinctsummarization, Geerts04tilingdatabases} on finding hyper-rectangle based covers for tables. In both these cases, the emphasis is on completely covering or 
summarizing the table, suffering from the same problems as traditional drill down in that the user may be presented with
too many results. The techniques in the former case may end up picking rare ``patterns'' if they have high confidence, and in the latter case only work for a restricted number of attributes (in their case, 4). \agp{check}
